---
title: "Conjoint Analysis"
author: "Matt Meister"
subtitle: "University of San Francisco"
include-in-header:
  - text: |
      <style>
      #title-slide .title {
        font-size: 3em;
        color: black;
      }
      </style>
format: 
  revealjs:
    self-contained: true
    scrollable: true
    theme: simple
    fontsize: 20pt
code-block-bg: true
code-block-border-left: "#31BAE9"
editor: source
execute: 
  error: true
  echo: true
  include: true
  warning: false
---

```{r, include=FALSE}
package.load <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    suppressMessages(install.packages(pkg, dependencies = TRUE))
    suppressMessages(library(pkg, character.only = TRUE))
  } else {
    suppressMessages(library(pkg, character.only = TRUE))
  }
}

pkgs <- c("dplyr", "ggplot2", "tidyr")

for (pkg in pkgs) {
  package.load(pkg)
}

```

## Overview

-   Conjoint analysis - intro & concepts
    -   Product design concepts
-   Conjoint analysis - steps to perform
    -   Identify a set of relevant product attributes
    -   Define reasonable levels for those attributes
    -   Create product profiles
    -   Obtain consumer preferences for profiles
    -   Analyze the data for each respondent
    -   Simulate market outcomes

# Conjoint Analysis Introduction

## Conjoint Analysis Introduction

Good product design begins with understanding customers

::: incremental

-   A market-oriented organization focuses on identifying and satisfying the needs and wants of its customers
-   Conjoint analysis is a powerful analytical tool that can aid managers in understanding and quantifying these needs and wants

:::

## Product design concepts

::: incremental

- A product can be considered as a bundle of attributes 
  - Attributes provide value to consumers 
  
- Each attribute can attain one or more levels 
  - Levels are specific values of attributes 

:::

## Product design concepts: Example

A computer can be described as:

|     |   Attribute   | Level 1 | Level 2 | Level 3 |
|:---:|:-------------:|:-------:|:-------:|:-------:|
|  1  | **Processor** | 2.4 GHz | 3.2 GHz |         |
|  2  |    **RAM**    |  8 GB   |  16 GB  |  32 GB  |
|  3  |  **Storage**  | 520 GB  | 1040 GB | 2080 GB |
|  4  |   **Price**   |  \$400  |  \$800  | \$1,200 |
|  5  |   **Color**   |  Grey   | Silver  |  Black  |


Sample profiles: 

::: incremental

- 2.4 GHz processor, 8 GB RAM, 520 GB hard drive, \$400 price, black 
- 2.4 GHz processor, 16 GB RAM, 520 GB hard drive, \$800 price, grey 
- 3.2 GHz processor, 8 GB RAM, 1040 GB, hard drive, \$800 price, silver 
- 3.2 GHz processor, 32 GB RAM, 1040 GB hard drive, \$1,200 price, silver

:::

## How do we know what product people will like?

Need a reliable method to measure preferences

::: incremental

- Two approaches:
  - Stated preference
    -Directly ask customers how much they value attributes
  - Revealed preference
    -Elicit customer preferences for whole products (profiles)
    -Infer value of attributes from responses

:::

## Advantages of revealed preference approach

::: incremental

- Easier for respondents
- Hard to answer questions like: 
  - "How much do you value a 4 hour battery life on your VR headset?"
- More accurate
  - Counters tendency to rate all attributes as (equally) important

:::

## Conjoint: Revealed preference via experiment

::: incremental

- Attributes are considered jointly
  - Subjects rate/rank/choose between product profiles
  - Subjects forced to make trade-offs for attribute levels
  - Trade-offs reveal true valuation
- Limitations
  - Attributes must be known in advance
  - Hard to handle a large (>10) number of attributes

::: 


## Stages in conjoint analysis

::: incremental

- Identify a set of relevant product attributes
- Define reasonable levels for those attributes
- Create product profiles
- Obtain consumer preferences for profiles
- Analyze the data for each respondent
- Simulate market outcomes

:::

## Stages 1 & 2: Attribute & level selection guidelines

Attributes in conjoint should be:

::: incremental

- Clear and unambiguous
- Actionable by the firm
- The total number of attributes should be kept low
  - 5-6 is the average, most studies fall between 4 and 8
- Levels should span the realistic range of possible values
  - E.g., include price levels close to min and max of market prices
  - Use qualitative research and pretests to decide on attributes and levels

:::

## Stages 1 & 2 example: VR Headset

- Standalone:
  - Yes vs no
- Cellular network access:
  - Yes vs no
- Price:
  - \$400 vs \$800 vs \$1,200
- Battery life:
  - 4 hours vs 8 hours vs 12 hours


## How many possible profiles are there

2 attributes with 3 levels and 2 with 2 levels?

::: fragment

$3 \times 3 \times 2 \times 2$

:::

::: fragment

$= 36$

:::

## Stage 3: Create product profiles

Data collection needs simplification (why?)

::: incremental

- Shrink total set to only include possible/realistic products
  - No impossible combinations
- Unrealistic products can ruin our data

:::

## Stage 4: Obtain consumer preferences for profiles

::: incremental

- Rating
  - Consumers see a set of profiles, rate each
    - **Positives?**
      - People are used to rating things
      - Simple
      - Straightforward
    - **Negatives?**
      - Not a choice! (Not economically relevant necessarily)
      - Ratings are not necessarily comparable across products
- Ranking
- Choice

:::

## Stage 4: Obtain consumer preferences for profiles

::: incremental

- Rating
- Ranking
  - People see all choices and rank them
    - **Positives?**
      - Simple
      - Straightforward
    - **Negatives?**
      - Still not a choice
      - What if things aren't evenly spaced?
- Choice

:::

## Stage 4: Obtain consumer preferences for profiles

::: incremental

- Rating
- Ranking
- Choice
  - People see options (usually 2+ at a time) and choose
    - **Positives?**
      - Real!
      - Potentially simple
    - **Negatives?**
      - Need a **lot** of trials
      
:::

## Stage 5: Data analysis

::: incremental

- Typically, conducted at the individual level
- Recovers a set of utility parameter estimates for each individual
- Mode of analysis dependent on response format:
  - Ratings data — linear regression (OLS)

:::

## Stage 5: Data analysis

Read in these responses to a conjoint survey for tablet computers

```{r}
responses_DF <- read.csv("respondent_data.csv") # survey reponses ("Y" variables)
N <- nrow(responses_DF) # number of subjects
summary(responses_DF)
```

## Stage 5: Data analysis

For tablet computers, a relevant set of product attributes might be:

1. Screen size (inches)
2. Cellular network connectivity
3. Price
4. Battery life (hrs)
5. Operating system (OS) 

## Stage 5: Data analysis

For tablet computers, a relevant set of product attributes might be:

| **Attribute**  | **1** | **2** | **3** | **4** | **5** |
| --- | --- | --- | --- | --- | --- |
| **Level**  | **Screen (in)** | **Cell** | **Price ($)** | **Battery (hr)** | **OS** |
| **1** | 7 | N | 100 | 4 | Android |
| **2** | 10 | Y | 300 | 8 |  iOS |
| **3** |   |   | 500 | 12 | Windows |

## Stage 5: Data analysis

For tablet computers, a relevant set of product **combinations** might be:

| **Profile #** | **Screen (in)** | **Cell** | **Price ($)** | **Battery (hr)** | **OS** |
| --- | --- | --- | --- | --- | --- |
| **1** | 7 | Y | 300 | 12 | iOS |
| **2** | 7 | Y | 100 | 8 | iOS |
| **3** | 10 | Y | 500 | 12 | Windows |
| **4** | 7 | Y | 300 | 4 | Windows |
| **5** | 7 | N | 300 | 8 | Android |
| **6** | 10 | N | 300 | 12 | iOS |
| **7** | 7 | N | 500 | 12 | Windows |
| **8** | 10 | N | 300 | 8 | Windows |
| **9** | 7 | N | 500 | 4 | iOS |
| **10** | 10 | Y | 100 | 12 | Android |
| **11** | 10 | Y | 300 | 4 | Android |
| **12** | 7 | N | 100 | 12 | Android |
| **13** | 10 | Y | 500 | 8 | iOS |
| **14** | 10 | N | 500 | 4 | Android |
| **15** | 10 | N | 100 | 4 | iOS |
| **16** | 10 | N | 100 | 8 | Windows |
| **17** | 7 | Y | 500 | 8 | Android |
| **18** | 7 | Y | 100 | 4 | Windows |

## Stage 5: Data analysis

```{r}
design_DF <- read.csv("survey_design.csv") # survey design ("X" variables)
design_DF
```

## Stage 5: Data analysis

Let's say we have a customer who responds to those profiles like this:

```{r}
response <- c(3,4,4,2,4,4,3,4,1,7,4,6,2,3,4,6,4,5)
```

::: fragment

```{r}
est_DF <- cbind(design_DF, response=response)

lm1 <- lm(response ~ factor(Screen) + factor(Cell) + factor(Price) + 
            factor(Battery) + factor(OS), 
          data=est_DF)

summary(lm1)
```

:::

## Stage 5: Data analysis

#### Ratings Data

::: incremental

- Intercept captures utility of "baseline" profile
  - Baseline profile corresponds to all omitted attribute levels
- Part-worths measure incremental utility relative to the baseline
  - Part-worths of included attribute levels = regression coefficients
  - Part-worths of omitted attribute levels = 0 by definition

:::

## Stage 5: Data analysis

#### Ratings Data

::: incremental

- *What is the baseline profile?*
  - The baseline profile is the profile associated with the omitted factor levels from the dummy variable coding. In this case, the baseline profile is a 7" screen, no cellular connectivity, $100 price, 4 hour battery life, and Android OS. 
- *What is the expected utility (rating) for the baseline profile?*
  - The utility/rating for this profile is given by the intercept estimate.
  - 4.3472
- *Interpret the coefficients from the regression*
  - Dummy variable coefficients represent in the incremental utility (rating), relative to the baseline profile
  - e.g., the coefficient on `factor(Screen)10` represents the incremental utility of changing from a 7" screen to a 10" screen
  
:::

## Step 5: Data analysis

Generate similar estimates for each individual represented in our supplemental data, `responses_DF`

::: fragment

How?

:::

::: incremental

- Create an empty list-of-lists to hold the regression results for each individual
  - `lm_res <- vector(mode = "list", length=nrow(responses_DF))`
- Loop over individuals. For each individual `i`:
  - Create a dataframe that combines design_DF with the responses for individual `i`.
- Estimate a linear model with individual `i`’s responses as the dependent variable, and the columns of design_DF as **(factor)** regressors
- Store the linear model results to the `i`’th element of lm_res. 
  - To access or assign (top level) lists in a list-of-lists, we must use double bracket indexing, as in: `lm_res[[i]]`

:::

## Step 5: Data analysis

```{r}
## estimate regression models, 1 per individual
# initialize an empty "list of lists" to hold lm() regression results
lm_res <- vector(mode="list", length=nrow(responses_DF))
# loop over subjects
for (i in 1:nrow(responses_DF)) {
  # get survey reponses for subject i (dropping respondent_id)
  response = as.numeric(responses_DF[i,2:ncol(responses_DF)])
  # create "estimation" dataframe, including i's reponses (Y) and the design variables (X)
  est_DF = cbind(design_DF, response=response)
  # run regression for subject i, store as lm_res[[i]]
  # note use of double bracket indexing [[ ]] syntax to access top-level lists
  lm_res[[i]] = lm(response ~ factor(Screen) + factor(Cell) + factor(Price) + 
                     factor(Battery) + factor(OS),
                   data=est_DF)
  }
```


## Stage 6: Simulate market outcomes

::: incremental

- Consider a hypothetical market
  - Start with closest approximation to current competitive landscape
  - Add 1st new product under consideration for introduction
- Compute expected utility for each of the products, for each subject in the sample
- Predict market shares (or units sold) for all products:
  - Assume subjects choose product with the highest utility
  - Product share = fraction of subjects who choose that product
- Repeat steps above for remaining new product designs
  - Introduce product concept with highest expected profit/market share

:::

## Stage 6

::: incremental

- Conjoint allows us to evaluate "what if" scenarios
  - How would a hypothetical "new" product would fare in competition with existing products?
- In the context of our example
  - How a new product by Toshiba will compete against the existing iPad
  - Assume the existing "iPad" product corresponds to:
    - Screen = 10 (inches)
    - Cell = "Y" (has cell connectivity)
    - Price = 500 ($)
    - Battery = 8 (hrs)
    - OS = "iOS"

:::

## Stage 6

Assume (initially) that Toshiba is considering one potential product design, `Toshiba_A`

::: incremental

- Screen = 7 (inches)
- Cell = "N" (no cell connectivity)
- Price = 300 ($)
- Battery = 12 (hrs)
- OS = "Android"
- The cost to produce a tablet computer is:
  - $(75 + 5\times(Screen==10) + 20\times(Cell==Y) + 5\times(Battery==8) + 15\times(Battery==12))$

:::

## Stage 6

Define `Toshiba_A` as the first product and the iPad as the second in a dataframe:

```{r}
prods1 <- data.frame(Screen = c(7,10),
                     Cell = c("N","Y"),
                     Price = c(300,500),
                     Battery = c(12,8),
                     OS = c("Android","iOS"))

rownames(prods1) = c("Toshiba_A","iPad")
prods1

```

## Stage 6

Use `lm1` to predict a choice between the iPad and `Toshiba_A`

*Hints:*

- Compute the utility (rating) that each consumer would derive from each alternative. You can use `predict()` for this task.
- Predict choice as the product with the highest utility (rating). The `which.max()` function can be useful here

::: fragment

```{r}
ratings = predict(lm1, newdata=prods1)
choice = which.max(ratings)
P = prods1[1,"Price"]
MC = (75 + 5*(prods1[1,"Screen"]==10) + 20*(prods1[1,"Cell"]=="Y") + 5*(prods1[1,"Battery"]==8) + 15*(prods1[1,"Battery"]==12))
profit = (choice==1)*(P-MC)
```

:::

## Stage 6

Discussion

::: incremental

- Which product did the model predict we would choose? 
  - iPad
- What is the cost of Toshiba_A?
  - Toshiba_A costs $75 + 15 = \$90$
  
:::

## Stage 6

Predict choices for supplemental survey respones

::: incremental

- Create a function called tablet_pft1 that takes as inputs:
  - `lm_res` containing regression results for all subjects
  - `prods`, a dataframe containing product definitions, as in the previous section
- `tablet_pft1` should return:
  - `profit` - the profit for Toshiba’s product (product 1)
  - `cost` - the cost of Toshiba’s product (product 1)
  - `choices` - predicted product choices for each subject in `responses_DF` (a list)

:::

## Stage 6

Predict choices for supplemental survey respones

```{r}
## function: tablet_pft1
# inputs:
# lm_res = list of lm() regression results
# prods_DF = dataframe of product definitions
# outputs:
# choices = predicted choice of each subject (list)
# cost = cost of product 1
# profit = profit of product 1
tablet_pft1 <- function(lm_res, prods_DF) {
  # initialize
  N = length(lm_res) # number subjects
  choices = rep(0,N)
  # loop over subjects: predict ratings/utilities, determine expected choices
  for (i in 1:N) {
    ratings = predict(lm_res[[i]], newdata=prods_DF)
    choices[i] = which.max(ratings)
    }
  # calculate demand, cost, profits for product 1 (1st row in prods_DF)
  Q = sum(choices==1)
  P = prods_DF[1,"Price"]
  MC = (75 + 5*(prods_DF[1,"Screen"]==10) + 20*(prods_DF[1,"Cell"]=="Y") + 
          5*(prods_DF[1,"Battery"]==8) + 15*(prods_DF[1,"Battery"]==12))
  pft = Q*(P-MC)
  # return values
  return(list(choices=choices, cost=MC, profit=pft))
  }
```

## Stage 6

Use tablet_pft1 to report:

- profits for `Toshiba_A`
- demand (expected units sold) for `Toshiba_A`
- the cost of `Toshiba_A`
- *Hints*
  - For each subject, compute the utility (rating) that each consumer would derive from each alternative
  - For each subject, predict choice as the product with the highest utility (rating)

```{r}
# call tablet_pft1 using prods1
prods1.results <- tablet_pft1(lm_res, prods1)
prods1.results$profit # profits
sum(prods1.results$choices==1) # units sold
prods1.results$cost # marginal cost
```

## Stage 6

Toshiba_B alternative product

::: incremental

- Assume Toshiba is considering between releasing the `Toshiba_A` product and `Toshiba_B`
  - Screen = 10 (inches)
  - Cell = "N" (no cell connectivity)
  - Price = 300 ($)
  - Battery = 12 (hrs)
  - OS = "Android"
- Repeat the analysis above, assuming the iPad competes against `Toshiba_B` only

:::

## Stage 6

Toshiba_B alternative product

```{r}
prods2 <- data.frame(Screen=c(10,10),
                    Cell=c("N","Y"),
                    Price=c(300,500),
                    Battery=c(12,8),
                    OS=c("Android","iOS"))
rownames(prods1) <- c("Toshiba_B","iPad")
```

## Stage 6

Toshiba_B alternative product

```{r}
prods2.results <- tablet_pft1(lm_res, prods2)
prods2.results$profit # profits
sum(prods2.results$choices==1) # units sold
prods2.results$cost # marginal cost
```

## Stage 6

Toshiba_B alternative product

::: incremental

- Which product should Toshiba introduce?
  - The profit from Toshiba_B is higher ($9635 vs $7350), so Toshiba should go with product B
  
:::

## Stage 6

Search over all Android OS alternative products

::: incremental

- The trick here is first to enumerate all possible Android-based tablet designs. The `expand.grid()` function is useful for this purpose
- After enumerating all possible Android-based tablet designs, loop over the designs and compute profits as before. Store the profit values in a list.
- Find the element of the profit list with the highest profit. Use the index value for this profile to print the associated product attribute levels.

:::

## Stage 6

```{r}
# create dataframe will all combinations of screen/cell/price/battery
allprods_DF <- expand.grid(unique(design_DF$Screen),
                           unique(design_DF$Cell),
                           unique(design_DF$Price),
                           unique(design_DF$Battery))
colnames(allprods_DF) <- c("Screen","Cell","Price","Battery")
```

## Stage 6

```{r}
Np <- nrow(allprods_DF) # number of products to test
# calculate profits for each candidate product, assuming iPad competition
pft <- rep(0,Np)
for (i in 1:Np) {
  prods = data.frame(Screen=c(allprods_DF[i,"Screen"],10),
                     Cell=c(as.character(allprods_DF[i,"Cell"]),"Y"),
                     Price=c(allprods_DF[i,"Price"],500),
                     Battery=c(allprods_DF[i,"Battery"],8),
                     OS=c("Android","iOS"))
  prods.results <- tablet_pft1(lm_res, prods)
  pft[i] <- prods.results$profit
}
```

## Stage 6

```{r}
# profit
max(pft)
# change from Toshiba_B
max(pft)-prods2.results$profit
100*(max(pft)-prods2.results$profit)/prods2.results$profit
# profit-maximizing product
allprods_DF[which.max(pft),]
```


## Guidelines

- Do not use too many attributes (<= 6)
  - More attributes increases the burden on consumers in at least two ways
    - Longer questionnaires 
    - Makes each question harder to answer, because products become harder to evaluate
- Focus on the attributes for which managerial decisions need to be made
- Do not use subjective attributes or ambiguous level descriptions
- Avoid infeasible combinations